---
title: "life_traits"
author: "Mireia Valle"
date: "8/9/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
##Specific libraries for getting functional traits
###FishLife
#devtools::install_github("DanOvando/FishLife", ref= "query-fishlife")
library(FishLife)

##Libraries for data management
library(tidyverse) 
library(data.table)
library(reshape)

##Library for correlation visualization
library(corrplot)

##Library for html tables
#install.packages ("kableExtra")
library (kableExtra)
```

# -------------------------------------------------------------------------------------------------------------------#
# LOADING AQUAMAPS SPECIES DATA 2018 
# -------------------------------------------------------------------------------------------------------------------#

## Loading species list 
```{r}
#Reading data from MAZU
dir_am <- '/home/shares/ohi/git-annex/globalprep/_raw_data/aquamaps'

spp <- data.table::fread(file.path(dir_am, 'd2018/speciesoccursum_ver0816c.csv'))

#From my laptop
#dir_am <- '/home/valle/github_annex/Aquamaps/'
#spp <- data.table::fread(file.path(dir_am, 'd2018/speciesoccursum_ver0816c.csv'))

#Tidy the species data table
head (spp)

spp <- spp %>%
  setNames(tolower(names(.))) %>%
  select(am_sid = speciesid, genus = genus, species = species, kingdom = kingdom, phylum = phylum, class = class, order = order, family = family)

colnames (spp)

summary (spp)
str(spp)
unique (spp$am_sid) %>% length()#24904
```

# -------------------------------------------------------------------------------------------------------------------#
# GETTING FUNCTIONAL TRAITS FROM FishLife
# -------------------------------------------------------------------------------------------------------------------#

## Notes on FishLife
### https://github.com/James-Thorson/FishLife
### http://onlinelibrary.wiley.com/doi/10.1002/eap.1606/full 
### * Loo = asymptotic maximum length (cm) 
### * K = individual growth rate (yr^-1) brody growth coefficient
### * Winfinity = asymptotic maximum mass (gr) 
### * tmax = maximum age (yr) 
### * tm = age at maturation (yr)
### * M = natural mortality rate (yr^-1)
### * Lm = length at maturation (cm) 
### * T = average temperature within distribution (deg C) 

## Function *Get_traits* that we are going to use was developed by Dan Ovando to query fishlife and store outputs in a sane structure, without this function the only way to extract life history parameters is by generating plots and then writing out their embedded predictions

```{r}
#?Get_traits
```

## For a dataframe with SEPARATE 'genus' and 'species' columns (e.g., 'Thunnus','thynnus'), run the following: 
```{r}
sfg = safely(Get_traits)

species_lt2018<- spp %>% 
  mutate(life_traits = map2(genus, species, ~(sfg(Genus=.x, Species=.y)))) %>%
  mutate(fish_life_worked = map(life_traits, 'error') %>% map_lgl(is.null)) %>%
  filter(fish_life_worked) %>%
  mutate(life_traits = map(life_traits, 'result')) %>%
  unnest() %>% 
  mutate(sciname = paste(genus, species))
```
## See results

```{r}
as.tibble (species_lt2018) 
head (species_lt2018)
colnames (species_lt2018)
str(species_lt2018)
unique (species_lt2018$am_sid) %>% length()#12368
unique (species_lt2018$fish_life_worked) 
```

## We subset the table with the columns we want to keep
```{r}
species_lt2018 <- subset(species_lt2018, select=c("am_sid", "genus", "species","sciname", "kingdom", "phylum", "class", "order", "family", "Loo", "K", "Winfinity", "tmax", "tm", "M", "Lm", "Temperature" ))
head (species_lt2018)
```

## Exploring the predicted values
```{r}
lt_predicted <- species_lt2018 %>% 
  group_by(order, Loo) %>% 
  mutate(unique_lt_values = n()) %>% 
  mutate(predicted_lt = ifelse (unique_lt_values > 1, "TRUE", "FALSE"))

lt_predicted_count <- lt_predicted %>% 
  group_by (predicted_lt) %>% 
  mutate(count_predicted_lt = n())

#predicted = 10730
#unique life traits values = 1638
```

## Visualizing the predicted values
```{r}
ggplot(lt_predicted , aes(order, unique_lt_values)) + 
  geom_point() +
  coord_flip()

lt_predicted_table <- lt_predicted %>% 
  group_by(family, unique_lt_values) %>% 
  select (order, family) %>% 
  distinct() %>% 
  arrange (desc(unique_lt_values))

kable(lt_predicted_table) %>% 
  kable_styling(bootstrap_options = "striped", 
                full_width = FALSE) 
```

## Analysing the data
```{r}
colnames (species_lt2018)
lt_matrix <- subset(species_lt2018, select=c("Loo", "K", "Winfinity", "tmax", "tm", "M", "Lm"))
head (lt_matrix)
```

## Analysing correlations: 
```{r}
cor_lt <- cor (lt_matrix) 
corrplot(cor_lt, method = "ellipse")
corrplot(cor_lt, method = "number") # Display the correlation coefficient
corrplot.mixed(cor_lt) # a wrapped function for mixed visualization style.
corrplot(cor_lt, type = "upper", order = "hclust")
plot (lt_matrix$Loo~lt_matrix$Lm)
```

## Save the results
```{r}
#write.csv (lt_predicted_count , file = "data/species_lt2018_predicted_count.csv")
```

