---
title: "traits_matrix"
author: "Mireia Valle"
date: "8/19/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r}
#installing and loading libraries
library (tidyverse)
library (here)
#install.packages ("naniar")
library(naniar)#to explore missing values
library(ggplot2)
#install.packages("skimr")
library(skimr)
library (kableExtra)
```

## Creating the traits matrix joining the life traits data frame with the fishbase dataset
```{r}
#setwd("~/github_annex")
```
## Reading the table we got from FishLife

```{r}
species_lt2018 <- read_csv(here("data", "species_lt2018_dropped.csv"))
unique (species_lt2018$am_sid) %>% length () #12363
sum(is.na(species_lt2018$am_sid))
```
## Reading the table we got from Fishbase

```{r}
fishbase_traits <- read_csv (here("data", "fishbase_traits.csv"))
as.tibble (fishbase_traits)
unique (fishbase_traits$sciname) %>% length () #12204
sum(is.na(fishbase_traits$sciname)) #0
```

## Reading the table from Beukhof et al 2019 https://doi.pangaea.de/10.1594/PANGAEA.900866 A trait collection of marine fish species from North Atlantic and Northeast Pacific continental shelf seas. Sourced: https://opentraits.org/datasets.html
```{r}
north_traits <- read_csv (here("data", "TraitCollectionFishNAtlanticNEPacificContShelf_Beukhofetal2019.csv"))
as.tibble (north_traits)
unique (north_traits$taxon) %>% length () #2010
sum(is.na(north_traits$taxon)) #138

# there are more than one value for the different traits, we need to summarize by spp in order to join this table to our other tables. We will do selecting the first value, but it will be interesting to explore other options

north_traits_spp <- north_traits %>%
  group_by(taxon)  %>%
  summarize(family  = first (family) , 
            genus = first (genus), 
            species = first (species),
            taxonomic.rank = first (taxonomic.rank), 
            LME = first (LME), 
            FAO = first (FAO), 
            habitat = first (habitat), 
            reference.habitat = first (reference.habitat),
            feeding.mode = first (feeding.mode),
            reference.feeding.mode = first (reference.feeding.mode), 
            tl = first (tl), 
            level.tl = first (level.tl), 
            reference.tl = first (reference.tl), 
            body.shape = first (body.shape), 
            fin.shape = first (fin.shape), 
            AR = first (AR), 
            level.AR = first (level.AR),
            reference.AR = first (reference.AR), 
            offspring.size = first (offspring.size), 
            level.offspring.size = first (level.offspring.size), 
            reference.offspring.size = first (reference.offspring.size), 
            spawning.type = first (spawning.type), 
            reference.spawning.type = first (reference.spawning.type),
            age.maturity = first (age.maturity), 
            level.age.maturity = first (level.age.maturity), 
            reference.age.maturity = first (reference.age.maturity), 
            fecundity = first (fecundity), 
            level.fecundity = first (level.fecundity), 
            reference.fecundity = first (reference.fecundity), 
            length.infinity = first (length.infinity), 
            growth.coefficient = first (growth.coefficient),
            level.growth = first (level.growth), 
            reference.growth = first (reference.growth), 
            length.max = first (length.max), 
            level.length.max = first (level.length.max), 
            reference.length.max = first (reference.length.max), 
            age.max = first (age.max),
            level.age.max = first (level.age.max), 
            reference.age.max = first (reference.age.max))

# Select the traits we want to keep 
north_traits_spp <- north_traits_spp  %>% 
  select (sciname = taxon,
          family  =  family , 
          genus =  genus, 
          species =  species,
          taxonomic.rank =  taxonomic.rank, 
          LME =  LME, 
          FAO =  FAO, 
          habitat =  habitat, 
          reference.habitat =  reference.habitat,
          feeding.mode =  feeding.mode,
          reference.feeding.mode =  reference.feeding.mode, 
          tl =  tl, 
          level.tl =  level.tl, 
          reference.tl =  reference.tl, 
          body.shape =  body.shape, 
          fin.shape =  fin.shape, 
          AR =  AR, 
          level.AR =  level.AR,
          reference.AR =  reference.AR, 
          offspring.size =  offspring.size, 
          level.offspring.size =  level.offspring.size, 
          reference.offspring.size =  reference.offspring.size, 
          spawning.type =  spawning.type, 
          reference.spawning.type =  reference.spawning.type,
          age.maturity =  age.maturity, 
          level.age.maturity =  level.age.maturity, 
          reference.age.maturity =  reference.age.maturity, 
          fecundity =  fecundity, 
          level.fecundity =  level.fecundity, 
          reference.fecundity =  reference.fecundity, 
          length.infinity =  length.infinity, 
          growth.coefficient =  growth.coefficient,
          level.growth =  level.growth, 
          reference.growth =  reference.growth, 
          length.max =  length.max, 
          level.length.max =  level.length.max, 
          reference.length.max =  reference.length.max, 
          age.max =  age.max,
          level.age.max =  level.age.max, 
          reference.age.max =  reference.age.max)
```

## We first join FishLife and Fishbase tables
```{r}
traits_matrix_lt_fishbase <- left_join (fishbase_traits, species_lt2018, by= "sciname")

as.tibble (traits_matrix_lt_fishbase)
unique (traits_matrix_lt_fishbase$sciname) %>% length () #12204
sum(is.na(traits_matrix_lt_fishbase$sciname)) # 0
```

## We join north traits table to the one we have created above

```{r}
traits_matrix <- left_join (traits_matrix_lt_fishbase, north_traits_spp, by= "sciname")

#we clean the data frame
traits_matrix <- traits_matrix %>% 
  select (-X1.x, -Species, -X1.y, -X1_1, -X, -family.y, -genus.y, -species.y)

#we save it
#write.csv (traits_matrix, file = "traits_matrix.csv")
```

## We explore the data 
```{r}
summary (traits_matrix)
colnames (traits_matrix)
skim <- skim(traits_matrix)

skim_table <- skim %>% 
  select (skim_variable, n_missing, complete_rate) %>% 
  arrange (desc(n_missing))

kable(skim_table) %>% 
  kable_styling(bootstrap_options = "striped", 
                full_width = FALSE) 

#Exploring NA's using naniar::gg_miss_var()

gg_miss_var(traits_matrix)

order_NAs <- traits_matrix %>%
  group_by(order) %>%
  miss_var_summary()


gg_miss_fct(x = traits_matrix, fct = class) + labs(title = "NA in traits matrix by class")

## We need to make decision on which var we want to keep and which we want to remove
#labels (traits_matrix)

#selected traits

traits_selected <- subset(traits_matrix, select=c(
  "SpecCode", 
  "sciname", 
  "am_sid", 
  "genus.x",
  "species.x",
  "kingdom", 
  "phylum", 
  "class", 
  "order", 
  "family.x", 
  "taxonomic.rank", 
  "Loo",
  "length.infinity", 
  "length.max",
  "level.length.max",
  "reference.length.max",
  "K", 
  "level.growth",	
  "reference.growth", 
  "growth.coefficient",
  "Winfinity", 
  "tmax", 
  "level.age.max",
  "reference.age.max", 
  "age.max", 
  "tm" , 
  "age.maturity",
  "level.age.maturity",
  "reference.age.maturity",
  "M" , 
  "Lm" , 
  "Temperature" ,
  "predicted_lt",
  "unique_lt_values",
  "count_predicted_lt",
  "bodyshape",
  "body.shape",
  "BodyShapeI",
  "demerspelag",
  "habitat",
  "reference.habitat",
  "depthrangedeep", 
  "vulnerability", 
  "importance", 
  "pricecateg", 
  "pd50", 
  "troph", 
  "setrop", 
  "tl",
  "level.tl",
  "reference.tl",
  "DietTroph", 
  "DietSeTroph", 
  "FoodTroph", 
  "FoodSeTroph", 
  "FeedingType", 
  "feeding.mode",
  "reference.feeding.mode",
  "fecuntype", 
  "fecun_min", 
  "fecun_max", 
  "fecundity", 
  "level.fecundity", 
  "reference.fecundity",
  "level.offspring.size",
  "reference.offspring.size",
  "offspring.size",
  "spawning.type",
  "reference.spawning.type",  
  "OperculumPresent", 
  "SexualAttributes",
  "SexMorphology", 
  "TypeofEyes", 
  "PosofMouth", 
  "TypeofMouth", 
  "CaudalFinI", 
  "CShape", 
  "fin.shape",
  "PelvicsAttributes", 
  "IUCN_Code", 
  "Protected", 
  "EnvTemp", 
  "Resilience", 
  "AdultType", 
  "AdultMode",
  "level.AR",
  "reference.AR",
  "AR",
  "LME",
  "FAO" ,
  "Epipelagic",
  "Mesopelagic",
  "Bathypelagic",
  "Abyssopelagic",
  "Hadopelagic",
  "Benthic",
  "Sessile",
  "Mobile",
  "Demersal",
  "Endofauna",
  "Pelagic"))
  

#we save the table
#write.csv (traits_selected, file = "traits_selected.csv")
```
## Exploring the data once we selected the traits
```{r}
## We explore the data 
summary (traits_selected)
colnames (traits_selected)
skim_selected <- skim(traits_selected)

skim_table_selected <- skim_selected %>% 
  select (skim_variable, n_missing, complete_rate) %>% 
  arrange (desc(n_missing))

kable(skim_table_selected) %>% 
  kable_styling(bootstrap_options = "striped", 
                full_width = FALSE) 

#Exploring NA's using naniar::gg_miss_var()

gg_miss_var(traits_selected)

order_NAs <- traits_selected %>%
  group_by(order) %>%
  miss_var_summary()


gg_miss_fct(x = traits_selected, fct = class) + labs(title = "NA in traits matrix by class")
```

## Wrangling the data for further analysis 

```{r}
traits_selected_short <- subset(traits_matrix, select=c("am_sid",
                                                  "class",
                                                  "order",
                                                  "family.x",
                                                  "demerspelag",
                                                  "bodyshape",
                                                  "IUCN_Code",
                                                  "Resilience",
                                                  "EnvTemp",
                                                  "FeedingType",
                                                  "importance",
                                                  "pricecateg",
                                                  "Loo",
                                                  "K",
                                                  "Winfinity",
                                                  "tmax",
                                                  "tm",
                                                  "M",
                                                  "Lm",
                                                  "depthrangedeep",
                                                  "pd50",
                                                  "vulnerability"))
```


```{r}
USE MUTATE
```


